# Assignment 1 Answers {.unnumbered}

## An Analytical Detective

There are two main types of crimes: violent crimes, and property crimes. In this problem, we'll focus on one specific type of property crime, called "motor vehicle theft" (sometimes referred to as grand theft auto). This is the act of stealing, or attempting to steal, a car. In this problem, we'll use some basic data analysis in R to understand the motor vehicle thefts in Chicago. 

Please download the file [mvtWeek1.csv](https://courses.edx.org/assets/courseware/v1/96f9b8f751467da3a4b8a5be33e32905/asset-v1:MITx+15.071x+2T2020+type@asset+block/mvtWeek1.csv) for this problem (do not open this file in any spreadsheet software before completing this problem because it might change the format of the Date field).

***Start:***\
Read the dataset mvtWeek1.csv into R, using the read.csv function, and call the data frame "mvt".

```{r}
mvt <- read.csv("mvtWeek1.csv")
```

**1.1: How many rows of data (observations) are in this dataset?**\
Answer: 191641
```{r}
nrow(mvt)
```

**1.2: How many variables are in this dataset?**\
Answer: 11
```{r}
ncol(mvt)
```

**1.3: Using the "max" function, what is the maximum value of the variable "ID"?**\
Answer: 9181151
```{r}
max(mvt$ID)
```

**1.4: What is the minimum value of the variable "Beat"?**\
Answer: 111
```{r}
min(mvt$Beat)
```

**1.5: How many observations have value TRUE in the Arrest variable (this is the number of crimes for which an arrest was made)?**\
Answer: 15536
```{r}
sum(mvt$Arrest)
```

**1.6: How many observations have a LocationDescription value of ALLEY?**\
Answer: 2308
```{r}
sum(mvt$LocationDescription == "ALLEY")
```

**2.1: In what format are the entries in the variable Date?**\
Answer: Month/Day/Year Hour:Minute
```{r}
mvt$Date[1]
```

**2.2: What is the month and year of the median date in our dataset? Enter your answer as "Month Year", without the quotes.**\
Answer: May 2006
```{r}
DateConvert = as.Date(strptime(mvt$Date, "%m/%d/%y %H:%M"))
#summary(DateConvert)
median(DateConvert)
```

**2.3: In which month did the fewest motor vehicle thefts occur?**\
Answer: February
```{r}
mvt$Month = months(DateConvert)
mvt$Weekday = weekdays(DateConvert)
mvt$Date = DateConvert
table(mvt$Month)
```

**2.4: On which weekday did the most motor vehicle thefts occur?**\
Answer: Friday
```{r}
table(mvt$Weekday)
```

**2.5: Which month has the largest number of motor vehicle thefts for which an arrest was made?**\
Answer: January
```{r}
table(mvt$Month, mvt$Arrest)
```

**3.1.1: In general, does it look like crime increases or decreases from 2002 - 2012?**\
Answer: Decreases
  
**3.1.2: In general, does it look like crime increases or decreases from 2005 - 2008?**\
Answer: Decreases
  
**3.1.3: In general, does it look like crime increases or decreases from 2009 - 2011?**\
Answer: Increases
```{r}
hist(mvt$Date, breaks=100)
```

**3.2: Does it look like there were more crimes for which arrests were made in the first half of the time period or the second half of the time period?**\
Answer: First half
```{r}
boxplot(Date ~ Arrest, data = mvt)
```

**3.3: For what proportion of motor vehicle thefts in 2001 was an arrest made?**\
Answer: 0.1041173
```{r}
table(mvt$Year, mvt$Arrest)
2152/(18517+2152)
```

**3.4: For what proportion of motor vehicle thefts in 2007 was an arrest made?**\
Answer: 0.08487395
```{r}
table(mvt$Year, mvt$Arrest)
1212/(13068+1212)
```

**3.5: For what proportion of motor vehicle thefts in 2012 was an arrest made?**\
Answer: 0.03902924
```{r}
table(mvt$Year, mvt$Arrest)
550/(13542+550)
```

**4.1: Which locations are the top five locations for motor vehicle thefts, excluding the "Other" category?**\
Answer: STREET, PARKING LOT/GARAGE(NON.RESID.), ALLEY, GAS STATION, DRIVEWAY - RESIDENTIAL
```{r}
sort(table(mvt$LocationDescription), decreasing = TRUE)
```

**Create a subset of your data, only taking observations for which the theft happened in one of these five locations, and call this new data set "Top5".**
```{r}
Top5 <- subset(mvt, mvt$LocationDescription == "STREET"
       | mvt$LocationDescription == "PARKING LOT/GARAGE(NON.RESID.)" 
       | mvt$LocationDescription == "ALLEY"
       | mvt$LocationDescription == "GAS STATION"
       | mvt$LocationDescription == "DRIVEWAY - RESIDENTIAL")
```


**4.2: How many observations are in Top5?**\
Answer: 177510
```{r}
nrow(Top5)
```

**4.3: One of the locations has a much higher arrest rate than the other locations. Which is it?**\
Answer: Gas Station (Check percentages)
```{r}
Top5$LocationDescription = factor(Top5$LocationDescription)
table(Top5$LocationDescription, Top5$Arrest)
```

**4.4: On which day of the week do the most motor vehicle thefts at gas stations happen?**\
Answer: Saturday
```{r}
table(Top5$LocationDescription == "GAS STATION", Top5$Weekday)
```

**4.5: On which day of the week do the fewest motor vehicle thefts in residential driveways happen?**\
Answer: Saturday
```{r}
table(Top5$LocationDescription == "DRIVEWAY - RESIDENTIAL", Top5$Weekday)
```

## Stock Dynamics

A stock market is where buyers and sellers trade shares of a company, and is one of the most popular ways for individuals and companies to invest money. The size of the world stock market  is now estimated to be in the trillions. The largest stock market in the world is the New York Stock Exchange (NYSE), located in New York City. About 2,800 companies are listed on the NYSE. In this problem, we'll look at the monthly stock prices of five of these companies: IBM, General Electric (GE), Procter and Gamble, Coca Cola, and Boeing. The data used in this problem comes from Infochimps.

Please download the following files: [IBMStock.csv](https://courses.edx.org/assets/courseware/v1/4fc08d10f171aacf2ef61c6b4b5bb4d8/asset-v1:MITx+15.071x+2T2020+type@asset+block/IBMStock.csv), [GEStock.csv](https://courses.edx.org/assets/courseware/v1/448b8be4693d913c2b5153be0c0e25d6/asset-v1:MITx+15.071x+2T2020+type@asset+block/GEStock.csv), [ProcterGambleStock.csv](https://courses.edx.org/assets/courseware/v1/bb6ed54230b5b2e29fb66819ed535da0/asset-v1:MITx+15.071x+2T2020+type@asset+block/ProcterGambleStock.csv), [CocaColaStock.csv](https://courses.edx.org/assets/courseware/v1/4c30fd7f4f55e537989ca13a6db36289/asset-v1:MITx+15.071x+2T2020+type@asset+block/CocaColaStock.csv), [BoeingStock.csv](https://courses.edx.org/assets/courseware/v1/2e8c9fb294db48e5a999c747b317722d/asset-v1:MITx+15.071x+2T2020+type@asset+block/BoeingStock.csv) (do not open these files in any spreadsheet software before completing this problem because it might change the format of the Date field).

**Start:**\
1. Read the datasets into R, using the read.csv function, and call the data frames "IBM", "GE", "ProcterGamble", "CocaCola", and "Boeing", respectively.
```{r}
IBM <- read.csv("IBMStock.csv")
GE <- read.csv("GEStock.csv")
ProcterGamble <- read.csv("ProcterGambleStock.csv")
CocaCola <- read.csv("CocaColaStock.csv")
Boeing <- read.csv("BoeingStock.csv")
```

2. Before working with these data sets, we need to convert the dates into a format that R can understand. Take a look at the structure of one of the datasets using the str function. Right now, the date variable is stored as a factor. We can convert this to a "Date" object in R by using the following five commands (one for each data set):
```{r}
IBM$Date = as.Date(IBM$Date, "%m/%d/%y")
GE$Date = as.Date(GE$Date, "%m/%d/%y")
CocaCola$Date = as.Date(CocaCola$Date, "%m/%d/%y")
ProcterGamble$Date = as.Date(ProcterGamble$Date, "%m/%d/%y")
Boeing$Date = as.Date(Boeing$Date, "%m/%d/%y")
```


**1.1: Our five datasets all have the same number of observations. How many observations are there in each data set?**\
Answer: 480
```{r}
nrow(IBM)
# According to the assignment, use: str(IBM)
# We only need to use the command for one of the datasets, since they all have the same number of obs.
```

**1.2: What is the earliest year in our datasets?**\
Answer: 1970
```{r}
min(IBM$Date)
# According to the assignment, use: summary(IBM$Date)
# Again, we only need to use the command for one of the datsets, since the observations started on the same date.
```

**1.3: What is the latest year in our datasets?**\
Answer: 2009
```{r}
max(IBM$Date)
# According to the assignment, use: summary(IBM$Date)
# Again, we only need to use the command for one of the datsets, since the observations ended on the same date.
```

**1.4: What is the mean stock price of IBM over this time period?**\
Answer: 144.375
```{r}
mean(IBM$StockPrice)
# According to the assignment, use: summary(IBM$StockPrice)
```

**1.5: **\
Answer: 9.293636 
```{r}
min(GE$StockPrice)
# According to the assignment, use: summary(GE$StockPrice)
```

**1.6: **\
Answer: 146.5843
```{r}
max(CocaCola$StockPrice)
# According to the assignment, use: summary(CocaCola$StockPrice)
```

**1.7: **\
Answer: 
```{r}
median(Boeing$StockPrice)
# According to the assignment, use: summary(Boeing$StockPrice)
```

**1.8: **\
Answer: 18.19414
```{r}
sd(ProcterGamble$StockPrice)
```
Side note: According to the assignment, questions 1.2 - 1.7 should've been solved using the summary function. However, I used commands that would give more accurate answer. Along with the commands I used, I also wrote how the assignment could be solved using the summary function.


**2.1.1: Around what year did Coca-Cola has its highest stock price in this time period?**\
Answer: 1973

**2.1.2: Around what year did Coca-Cola has its lowest stock price in this time period?**\
Answer: 1980
```{r}
plot(CocaCola$Date, CocaCola$StockPrice, "l")
```

**2.2: In March of 2000, the technology bubble burst, and a stock market crash occurred. According to this plot, which company's stock dropped more?**\
Answer: Procter and Gamble
```{r}
plot(CocaCola$Date, CocaCola$StockPrice, "l", col = "red")
lines(ProcterGamble$Date, ProcterGamble$StockPrice, col = "blue")
abline(v=as.Date(c("2000-03-01")), lwd=2)
```

**2.3.1: Around 1983, the stock for one of these companies (Coca-Cola or Procter and Gamble) was going up, while the other was going down. Which one was going up?**\
Answer: CocaCola

**2.3.1: In the time period shown in the plot, which stock generally has lower values?**\
Answer: CocaCola
```{r}
plot(CocaCola$Date, CocaCola$StockPrice, "l", col = "red")
lines(ProcterGamble$Date, ProcterGamble$StockPrice, col = "blue")
```

