# Assignment 1 Answers {.unnumbered}

## An Analytical Detective

There are two main types of crimes: violent crimes, and property crimes. In this problem, we'll focus on one specific type of property crime, called "motor vehicle theft" (sometimes referred to as grand theft auto). This is the act of stealing, or attempting to steal, a car. In this problem, we'll use some basic data analysis in R to understand the motor vehicle thefts in Chicago.

To download this data, click: [mvtWeek1.csv](https://courses.edx.org/assets/courseware/v1/96f9b8f751467da3a4b8a5be33e32905/asset-v1:MITx+15.071x+2T2020+type@asset+block/mvtWeek1.csv)

***Start:***\
Read the dataset mvtWeek1.csv into R, using the read.csv function, and call the data frame "mvt".

```{r}
mvt <- read.csv("mvtWeek1.csv")
```

**1.1: How many rows of data (observations) are in this dataset?**\
Answer: 191641
```{r}
nrow(mvt)
```

**1.2: How many variables are in this dataset?**\
Answer: 11
```{r}
ncol(mvt)
```

**1.3: Using the "max" function, what is the maximum value of the variable "ID"?**\
Answer: 9181151
```{r}
max(mvt$ID)
```

**1.4: What is the minimum value of the variable "Beat"?**\
Answer: 111
```{r}
min(mvt$Beat)
```

**1.5: How many observations have value TRUE in the Arrest variable (this is the number of crimes for which an arrest was made)?**\
Answer: 15536
```{r}
sum(mvt$Arrest)
```

**1.6: How many observations have a LocationDescription value of ALLEY?**\
Answer: 2308
```{r}
sum(mvt$LocationDescription == "ALLEY")
```

**2.1: In what format are the entries in the variable Date?**\
Answer: Month/Day/Year Hour:Minute
```{r}
mvt$Date[1]
```

**2.2: What is the month and year of the median date in our dataset? Enter your answer as "Month Year", without the quotes.**\
Answer: May 2006
```{r}
DateConvert = as.Date(strptime(mvt$Date, "%m/%d/%y %H:%M"))
#summary(DateConvert)
median(DateConvert)
```

**2.3: In which month did the fewest motor vehicle thefts occur?**\
Answer: February
```{r}
mvt$Month = months(DateConvert)
mvt$Weekday = weekdays(DateConvert)
mvt$Date = DateConvert
table(mvt$Month)
```

**2.4: On which weekday did the most motor vehicle thefts occur?**\
Answer: Friday
```{r}
table(mvt$Weekday)
```

**2.5: Which month has the largest number of motor vehicle thefts for which an arrest was made?**\
Answer: January
```{r}
table(mvt$Month, mvt$Arrest)
```

**3.1.1: In general, does it look like crime increases or decreases from 2002 - 2012?**\
Answer: Decreases
  
**3.1.2: In general, does it look like crime increases or decreases from 2005 - 2008?**\
Answer: Decreases
  
**3.1.3: In general, does it look like crime increases or decreases from 2009 - 2011?**\
Answer: Increases
```{r}
hist(mvt$Date, breaks=100)
```

**3.2: Does it look like there were more crimes for which arrests were made in the first half of the time period or the second half of the time period?**\
Answer: First half
```{r}
boxplot(Date ~ Arrest, data = mvt)
```

**3.3: For what proportion of motor vehicle thefts in 2001 was an arrest made?**\
Answer: 0.1041173
```{r}
table(mvt$Year, mvt$Arrest)
2152/(18517+2152)
```

**3.4: For what proportion of motor vehicle thefts in 2007 was an arrest made?**\
Answer: 0.08487395
```{r}
table(mvt$Year, mvt$Arrest)
1212/(13068+1212)
```

**3.5: For what proportion of motor vehicle thefts in 2012 was an arrest made?**\
Answer: 0.03902924
```{r}
table(mvt$Year, mvt$Arrest)
550/(13542+550)
```

**4.1: Which locations are the top five locations for motor vehicle thefts, excluding the "Other" category?**\
Answer: STREET, PARKING LOT/GARAGE(NON.RESID.), ALLEY, GAS STATION, DRIVEWAY - RESIDENTIAL
```{r}
sort(table(mvt$LocationDescription))
```

**Create a subset of your data, only taking observations for which the theft happened in one of these five locations, and call this new data set "Top5".**
```{r}
Top5 <- subset(mvt, mvt$LocationDescription == "STREET"
       | mvt$LocationDescription == "PARKING LOT/GARAGE(NON.RESID.)" 
       | mvt$LocationDescription == "ALLEY"
       | mvt$LocationDescription == "GAS STATION"
       | mvt$LocationDescription == "DRIVEWAY - RESIDENTIAL")
```


**4.2: How many observations are in Top5?**\
Answer: 177510
```{r}
nrow(Top5)
```

**4.3: One of the locations has a much higher arrest rate than the other locations. Which is it?**\
Answer: Gas Station (Check percentages)
```{r}
Top5$LocationDescription = factor(Top5$LocationDescription)
table(Top5$LocationDescription, Top5$Arrest)
```

**4.4: On which day of the week do the most motor vehicle thefts at gas stations happen?**\
Answer: Saturday
```{r}
table(Top5$LocationDescription == "GAS STATION", Top5$Weekday)
```

**4.5: On which day of the week do the fewest motor vehicle thefts in residential driveways happen?**\
Answer: Saturday
```{r}
table(Top5$LocationDescription == "DRIVEWAY - RESIDENTIAL", Top5$Weekday)
```