# The Analytics Edge: Assignment 2 {.unnumbered}

The following link will lead you to the assignment on the edX website: [https://learning.edx.org/course/course-v1:MITx+15.071x+2T2020/block-v1:MITx+15.071x+2T2020+type\@sequential+block\@60d93a44280348d7a0a16663f92af0f7](https://learning.edx.org/course/course-v1:MITx+15.071x+2T2020/block-v1:MITx+15.071x+2T2020+type@sequential+block@60d93a44280348d7a0a16663f92af0f7)

## Climate Change

There have been many studies documenting that the average global temperature has been increasing over the last century. The consequences of a continued rise in global temperature will be dire. Rising sea levels and an increased frequency of extreme weather events will affect billions of people.

In this problem, we will attempt to study the relationship between average global temperature and several other factors.

The file [climate_change.csv](https://courses.edx.org/asset-v1:MITx+15.071x+2T2017+type@asset+block@climate_change.csv) contains climate data from May 1983 to December 2008.

***Start:***\
We are interested in how changes in these variables affect future temperatures, as well as how well these variables explain temperature changes so far. To do this, first read the dataset climate_change.csv into R.

```{r}
climateChange <- read.csv("/cloud/project/analyticsedge/Datasets/DatasetsUnit2/climate_change.csv")
```

Then, split the data into a *training set*, consisting of all the observations up to and including 2006, and a *testing set* consisting of the remaining years (hint: use subset). A training set refers to the data that will be used to build the model (this is the data we give to the lm() function), and a testing set refers to the data we will use to test our predictive ability.

```{r}
climateTrain <- subset(climateChange, Year <=2006)
climateTest <- subset(climateChange, Year > 2006)
```

Next, build a linear regression model to predict the dependent variable Temp, using MEI, CO2, CH4, N2O, CFC.11, CFC.12, TSI, and Aerosols as independent variables (Year and Month should NOT be used in the model). Use the training set to build the model.

```{r}
climateModel1 <- lm(Temp ~ MEI + CO2 + CH4 + N2O + CFC.11 + CFC.12 + TSI + Aerosols, data = climateTrain)
```

**1.1: What is the model R2 (the "Multiple R-squared" value)?**\
Answer: 0.7509

**1.2: Which variables are significant in the model?**\
Answer: MEI, CO2, CFC.11, CFC.12, TSI, Aerosols

```{r}
summary(climateModel1)
```

Current scientific opinion is that nitrous oxide and CFC-11 are greenhouse gases: gases that are able to trap heat from the sun and contribute to the heating of the Earth. However, the regression coefficients of both the N2O and CFC-11 variables are *negative*, indicating that increasing atmospheric concentrations of either of these two compounds is associated with lower global temperatures.

**2.1: Which of the following is the simplest correct explanation for this contradiction?**\
Answer: All of the gas concentration variables reflect human development - N2O and CFC.11 are correlated with other variables in the data set.

**2.2.1: Which of the following independent variables is N2O highly correlated with (absolute correlation greater than 0.7)?**\
Answer: CO2, CH4, CFC.12

**2.2.2: Which of the following independent variables is CFC.11 highly correlated with?**\
Answer: CH4, CFC.12

```{r}
cor(climateTrain)
```

Given that the correlations are so high, let us focus on the N2O variable and build a model with only MEI, TSI, Aerosols and N2O as independent variables. Remember to use the training set to build the model.

```{r}
climateModel2 <- lm(Temp ~ MEI + N2O + TSI + Aerosols, data = climateTrain)
```

**3.1: What is the coefficient of N2O in this reduced model?**\
Answer: 2.532e-02 (0.02532)

**3.2: What is the model R^2?**\
Answer: 0.7261

```{r}
summary(climateModel2)
```



```{r}
climateModel <- step(climateModel1)
```

**4.1: What is the R^2 value of the model produced by the step function?**\
Answer: 0.7508

**4.2: Which of the variable(s) were eliminated from the full model by the step function?**\
Answer: CH4

```{r}
summary(climateModel)
```

**5: Using the model produced from the step function, calculate temperature predictions for the testing data set, using the predict function. What is the testing set R^2?**\
Answer: 0.6286051
```{r}
predictTemp <- predict(climateModel, newdata = climateTest)
SSE = sum((predictTemp - climateTest$Temp)^2)
SST = sum((mean(climateTrain$Temp) - climateTest$Temp)^2)
R2 = 1 - SSE/SST
R2
```

